# Forest Recovery Report: R Notebook Skeleton

---
title: "Forest Structure & Composition Recovery Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load Libraries
```{r load-libs}
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia)
library(vegan)
library(RColorBrewer)
```

# Load Data & set theme
```{r load-data}

theme_set(theme_classic2(base_size = 10) +
            theme(axis.title = element_text(size = 10),
                  axis.text  = element_text(size = 10)))


dat_overlap <- fread("outData/full_table_overlap_23_25.csv")
```

# Summary: Sites & Structure
```{r summary-sites}
dat_master_subplot <- dat_overlap %>% 
  dplyr::select(plot, subplot, year) %>% 
  distinct()

n_plots_total <- n_distinct(dat_master_subplot$plot)
n_subplots_total <- n_distinct(dat_master_subplot$subplot)

cat("Total plots:", n_plots_total, "\n")
cat("Total subplots:", n_subplots_total)

```

## investigate indiviidual subplots 

Is it correct if CV = 0 if a single tree species has several stems?

```{r investigate_cubplots, echo=FALSE}

# Identify subplots where the same species appears in multiple vegtypes
df_distinct_vegtypes <- dat_overlap %>%
  filter(n>0) %>% 
  group_by(subplot, species) %>%
  summarize(n_vegtypes = n_distinct(vegtype), .groups = "drop") %>%
  filter(n_vegtypes > 1)

# Extract just the subplot IDs
unique_subplots_two_layers <- unique(df_distinct_vegtypes$subplot) # 110 from 1250~ 9%

# Optional: View full records of the overlaps
overlap_records <- dat_overlap %>%
  filter(n>0) %>%
  filter(subplot %in% unique_subplots_two_layers) %>%
  arrange(subplot, species, vegtype) %>% 
  select(subplot, species, vegtype,     hgt,     n)

# Print the results
length(unique_subplots_two_layers)

cat("Total # subplots with a one tree species over two vertical layers:", length(unique_subplots_two_layers))

```

# Stem Counts & Species Richness
```{r stem-richness}
 field_sub_summ <- dat_overlap %>%
   mutate(n = coalesce(n, 0L)) %>%
  group_by(plot, subplot, year, time_snc_full_disturbance) %>%
   summarise(
     stems_total = sum(n, na.rm = TRUE),
     sp_richness = if (stems_total > 0) n_distinct(species[n > 0]) else 0,
     shannon_sp  = if (stems_total > 0) vegan::diversity(n, index = "shannon") else 0,
     evenness_sp = if (sp_richness > 1) shannon_sp / log(sp_richness) else 0,
     effective_numbers = ifelse(stems_total > 0, exp(shannon_sp), NA_real_),
     mean_hgt = if (stems_total > 0) weighted.mean(hgt_est[n > 0], n[n > 0]) else NA_real_,
     .groups = "drop"
   )

head(field_sub_summ)
```

#  Trait CWMs
```{r trait-cwms}
wvar <- "n"
wfun <- function(x) ifelse(is.na(x) | x < 0, 0, x)

cwm_subplot <- dat_overlap %>%
  mutate(w = wfun(.data[[wvar]])) %>%
  filter(w > 0) %>%
  group_by(plot, subplot, year, time_snc_full_disturbance) %>%
  summarise(
    stems_with_traits = sum(w[!is.na(Shade_tolerance) & !is.na(Drought_tolerance)], na.rm = TRUE),
    stems_total       = sum(w, na.rm = TRUE),
    CWM_shade   = ifelse(stems_with_traits > 0,
                         sum(w * Shade_tolerance, na.rm = TRUE) / stems_with_traits, NA_real_),
    CWM_drought = ifelse(stems_with_traits > 0,
                         sum(w * Drought_tolerance, na.rm = TRUE) / stems_with_traits, NA_real_),
    .groups = "drop"
  )
```


# GAM Models
```{r gam-models}
# m_cv_tw <- gam(
#   cv_hgt ~ level + 
#     s(time_snc_full_disturbance, by = level, k = 5, bs = "cs") +
#     s(plot_id, bs = "re"),
#   data = filter(both_levels_re2, cv_hgt > 0)
# )
# 
# summary(m_cv_tw)
```

#  Seral Stages
```{r seral-stage-share}
# Insert stacked barplot code for early vs late
```

# Summary Tables
```{r summary-tables}
# subplot_summary_tbl <- field_sub_summ %>%
#   ungroup() %>%
#   select(year, mean_hgt, cv_hgt, shannon_sp, sp_richness) %>%
#   pivot_longer(-year, names_to = "metric", values_to = "value") %>%
#   group_by(metric, year) %>%
#   summarise(
#     n = n(),
#     mean = mean(value, na.rm = TRUE),
#     sd = sd(value, na.rm = TRUE),
#     median = median(value, na.rm = TRUE),
#     IQR = IQR(value, na.rm = TRUE),
#     .groups = "drop"
#   )
```

#  Visuals: Combined Panel Plots
```{r panel-plots}
#ggarrange(p_cv_hgt, p_mean_hgt, common.legend = TRUE)
```

# 🧩 Appendix: Species Stats
```{r appendix-species}
# Insert code for species totals, stem share, and occurrence
```

# 📎 Notes
- CV uses Bessel correction (n - 1).
- Plots are faceted by metric and scaled freely.
- Trait CWMs use stem counts as weights.
